Towards Patient-Specific Neural Field Modeling using Micro-Electrode Data

### 2011-05-04 ltx/data_paper/DataPaper.tex
edits to derivation of Xi4 in supporting informatiom
edits to product of two Gaussians derivation
added figures to data paper

Having approximated $q\left(\mathbf x_t\right)$ in the E-step, by substituting  \ref{eq:qTaylor} into equations \ref{eq:Xi3}-\ref{eq:Xi5} an approximate to the non-linear $\boldsymbol\Xi$-variables can be calculated which provides an approximate lower bound on the log-likelihood function (see supporting information), 
\begin{align}
\boldsymbol\Xi_{3}&\approx\sum_{t=0}^{T-1}\left[ \mathbf{\hat x}_{t+1}^\top\tilde{\boldsymbol\Sigma}_e^{-1}q(\mathbf{\hat x}_t)+\int_\Omega\boldsymbol \phi^\top(\mathbf r') \mathbf M_{t+1} \tilde{\boldsymbol\Sigma}_e^{-1} \boldsymbol{\Psi}(\mathbf{r}') f'(\boldsymbol \phi^\top(\mathbf r')\mathbf {\hat x}_t) d\mathbf{r}'\right] \\	
\boldsymbol\Xi_{4}&\approx\sum_{t=0}^{T-1}\left[q^\top(\mathbf{\hat x}_t)\tilde{\boldsymbol\Sigma}_e^{-1}q(\mathbf{\hat x}_t)+\Lambda_t\right] \\
 \boldsymbol\Xi_{5}&\approx\sum_{t=0}^{T-1}\left[ \mathbf{\hat x}_{t}^\top\tilde{\boldsymbol\Sigma}_e^{-1}q(\mathbf{\hat x}_t)+\int_\Omega\boldsymbol \phi^\top(\mathbf r') \mathbf P_t \tilde{\boldsymbol\Sigma}_e^{-1}  \boldsymbol{\Psi}(\mathbf{r}') f'(\boldsymbol \phi^\top(\mathbf r')\mathbf {\hat x}_t) d\mathbf{r}'\right] 
 \end{align}
where
\begin{align}
	{\boldsymbol \lambda}_t^{(i)}&=\int_{\Omega} \boldsymbol \Psi(\mathbf r')\phi_i(\mathbf r')f'(\boldsymbol\phi^\top(\mathbf r')\mathbf {\hat x}_t)d\mathbf r' \label{eq:lambdatildei}
\end{align}
\begin{equation}
 \Lambda_t=\sum_{i,j=1}^{n_x}[\mathbf{P_t}]_{ij}{\boldsymbol\lambda}_t^{(i)\top}\tilde{\boldsymbol\Sigma}_e^{-1}{\boldsymbol\lambda}_t^{(j)}
\end{equation}

@@

% % all figures here
% % ~~~~~~~~~~~~~~~~
% \begin{figure}[!ht]
% \begin{center}
% \includegraphics{./Figures/LFPs.eps}
% \end{center}
% \caption{{\bf Local field potentials from the micro-electrode array}. Data recorded from the first 8 channels of the array. The red dotted line indicates the seizure onset and the blue dashed line indicates the seizure end.}
% \label{fig:TimeSeries}
% \end{figure}
% 
% 
% \begin{figure}[!ht]
% \begin{center}
% \includegraphics{./Figures/FieldObservations.eps}
% \end{center}
% \caption{{\bf Snap-shots of the spatial aspects of the observed field}. The snap-shots are ordered from A to I forming a consecutive sequence spaced 2.5~ms apart starting at 34.024s}
% \label{fig:FieldObserations}
% \end{figure}
% 
% \begin{figure}[!ht]
% \begin{center}
% \includegraphics{./Figures/TemporalFreq.eps}
% \end{center}
% \caption{{\bf Spectra from local field potential recordings}. A. Pre-seizure observations. B. Seizure observations. C. Post-seizure observations.}
% \label{fig:TemporalFreqObservation}
% \end{figure}
% 
% \begin{figure}[!ht]
% \begin{center}
% \includegraphics{./Figures/SpatialFreq.eps}
% \end{center}
% \caption{{\bf Spatial frequency analysis of the observed neural field}. A. Pre-seizure observations. B. Seizure observations. C. Post-seizure observations.}
% \label{fig:SpatialFreqObservation}
% \end{figure}
% 
% \begin{figure}[!ht]
% \begin{center}
% \includegraphics{./Figures/SpatialFreqCrossSection.eps}
% \end{center}
% \caption{{\bf Diagonal cross-section of the spatial frequency plots of the observed neural field}. A. Pre-seizure observations. B. Seizure observations. C. Post-seizure observations.}
% \label{fig:DiagSpatialFreqObservation}
% \end{figure}
% 
% \begin{figure}[!ht]
% \begin{center}
% \includegraphics{./Figures/CrossCorr2D.eps}
% \end{center}
% \caption{{\bf Mean cross-correlations between consecutive observations of the neural field}. Note: The correlations are normalised such that the maximum value is one. A. Pre-seizure observation correlations. B. Seizure observation correlations. C. Post-seizure observation correlations.}
% \label{fig:SpatialCrossCorrelation}
% \end{figure}
% 
% \begin{figure}[!ht]
% \begin{center}
% \includegraphics{./Figures/HomoTestCrossCorr.eps}
% \end{center}
% \caption{{\bf Mean cross-correlations between consecutive samples using subsets of observations of the neural field}. Each subplot shows correlations corresponding to a subset electrodes. Indexing the electrodes in a (x,y) grid, where index ranges from 1 to 10. Subplot A shows the mean correlation between electrode indexes (1-6, 5-10), subplot B shows correlations from indexes (5-10,5-10), subplot C shows indexes (1-6, 1-6) and subplot D shows indexes (5-10, 5-10).}
% \label{fig:HomogeneityTest}
% \end{figure}

>>>>

Having approximated $q\left(\mathbf x_t\right)$ in the E-step, by substituting  \ref{eq:qTaylor} into equations \ref{eq:Xi3}-\ref{eq:Xi5} an approximate to the non-linear $\boldsymbol\Xi$-variables can be calculated which provides an approximate lower bound on the log-likelihood function (see supporting information), 
\begin{align}
\boldsymbol\Xi_{3}&\approx\sum_{t=0}^{T-1}\left[ \mathbf{\hat x}_{t+1}^\top\tilde{\boldsymbol\Sigma}_e^{-1}q(\mathbf{\hat x}_t)+\int_\Omega\boldsymbol \phi^\top(\mathbf r') \mathbf M_{t+1} \tilde{\boldsymbol\Sigma}_e^{-1} \boldsymbol{\Psi}(\mathbf{r}') f'(\boldsymbol \phi^\top(\mathbf r')\mathbf {\hat x}_t) d\mathbf{r}'\right] \\	
\boldsymbol\Xi_{4}&\approx\sum_{t=0}^{T-1}\left[q^\top(\mathbf{\hat x}_t)\tilde{\boldsymbol\Sigma}_e^{-1}q(\mathbf{\hat x}_t)+\Lambda_t\right] \\
 \boldsymbol\Xi_{5}&\approx\sum_{t=0}^{T-1}\left[ \mathbf{\hat x}_{t}^\top\tilde{\boldsymbol\Sigma}_e^{-1}q(\mathbf{\hat x}_t)+\int_\Omega\boldsymbol \phi^\top(\mathbf r') \mathbf P_t \tilde{\boldsymbol\Sigma}_e^{-1}  \boldsymbol{\Psi}(\mathbf{r}') f'(\boldsymbol \phi^\top(\mathbf r')\mathbf {\hat x}_t) d\mathbf{r}'\right] 
 \end{align}
where
\begin{align}\label{eq:lambdat} 
	{\boldsymbol \lambda}_t^{(i)}&=\int_{\Omega} \boldsymbol \Psi(\mathbf r')\phi_i(\mathbf r')f'(\boldsymbol\phi^\top(\mathbf r')\mathbf {\hat x}_t)d\mathbf r'
\end{align}
and
\begin{equation}
 \boldsymbol\Lambda_t=\sum_{i,j=1}^{n_x}[\mathbf{P_t}]_{ij}{\boldsymbol\lambda}_t^{(i)\top}\tilde{\boldsymbol\Sigma}_e^{-1}{\boldsymbol\lambda}_t^{(j)}
\end{equation}

@@

% % all figures here
% % ~~~~~~~~~~~~~~~~
\begin{figure}[!ht]
\begin{center}
\includegraphics{./Figures/KernelWidthEstimation.pdf}
\end{center}
\caption{{\bf Connectivity kernel support estimation}.}
\label{fig:KernelWidthEstimation}
\end{figure}

\begin{figure}[!ht]
\begin{center}
\includegraphics{./Figures/KernelBasis.pdf}
\end{center}
\caption{{\bf Connectivity kernel basis functions, $n_{\theta}=25$}.}
\label{fig:KernelBasis}
\end{figure}
\begin{figure}[!ht]
\begin{center}
\includegraphics{./Figures/DisturbanceWidthEstimation.pdf}
\end{center}
\caption{{\bf Disturbance support estimation}.}
\label{fig:DisturbanceWidthEstimation}
\end{figure}

\begin{figure}[!ht]
\begin{center}
\includegraphics{./Figures/KernelEstimate.pdf}
\end{center}
\caption{{\bf Connectivity kernel estimates over 100 realisations.} left: True kernel; right;Estimated kernel.}
\label{fig:KernelEstimates}
\end{figure}

\begin{figure}[!ht]
\begin{center}
\includegraphics{./Figures/KernelEstimateCrossSection.pdf}
\end{center}
\caption{{\bf Cross-section of connectivity kernel estimates over 100 realisation.} True is blue; Estimate is green; 95\% Confedence interval is red.}
\label{fig:KernelEstimateCrossSection}
\end{figure}

\begin{figure}[!ht]
\begin{center}
\includegraphics{./Figures/ConfedenceIntervel.pdf}
\end{center}
\caption{{\bf Cross-section of connectivity kernel estimates over 100 realisation}. An alternative figure which shows estimates outside confedence interval.}
\label{figConfedenceIntervel}
\end{figure}

\begin{figure}[!ht]
\begin{center}
\includegraphics{./Figures/Histogram.pdf}
\end{center}
\caption{{\bf Synaptic dynamic parameter and hyperparameter estimates}.}
\label{fig:Histogram}
\end{figure}

\begin{figure}[!ht]
\begin{center}
\includegraphics{./Figures/Q.pdf}
\end{center}
\caption{{\bf Lower-bound of the log-likelihood}.}
\label{fig:FieldObserations}
\end{figure}

% \begin{figure}[!ht]
% \begin{center}
% \includegraphics{./Figures/LFPs.eps}
% \end{center}
% \caption{{\bf Local field potentials from the micro-electrode array}. Data recorded from the first 8 channels of the array. The red dotted line indicates the seizure onset and the blue dashed line indicates the seizure end.}
% \label{fig:TimeSeries}
% \end{figure}
% 
% 
% \begin{figure}[!ht]
% \begin{center}
% \includegraphics{./Figures/FieldObservations.eps}
% \end{center}
% \caption{{\bf Snap-shots of the spatial aspects of the observed field}. The snap-shots are ordered from A to I forming a consecutive sequence spaced 2.5~ms apart starting at 34.024s}
% \label{fig:FieldObserations}
% \end{figure}
% 
% \begin{figure}[!ht]
% \begin{center}
% \includegraphics{./Figures/TemporalFreq.eps}
% \end{center}
% \caption{{\bf Spectra from local field potential recordings}. A. Pre-seizure observations. B. Seizure observations. C. Post-seizure observations.}
% \label{fig:TemporalFreqObservation}
% \end{figure}
% 
% \begin{figure}[!ht]
% \begin{center}
% \includegraphics{./Figures/SpatialFreq.eps}
% \end{center}
% \caption{{\bf Spatial frequency analysis of the observed neural field}. A. Pre-seizure observations. B. Seizure observations. C. Post-seizure observations.}
% \label{fig:SpatialFreqObservation}
% \end{figure}
% 
% \begin{figure}[!ht]
% \begin{center}
% \includegraphics{./Figures/SpatialFreqCrossSection.eps}
% \end{center}
% \caption{{\bf Diagonal cross-section of the spatial frequency plots of the observed neural field}. A. Pre-seizure observations. B. Seizure observations. C. Post-seizure observations.}
% \label{fig:DiagSpatialFreqObservation}
% \end{figure}
% 
% \begin{figure}[!ht]
% \begin{center}
% \includegraphics{./Figures/CrossCorr2D.eps}
% \end{center}
% \caption{{\bf Mean cross-correlations between consecutive observations of the neural field}. Note: The correlations are normalised such that the maximum value is one. A. Pre-seizure observation correlations. B. Seizure observation correlations. C. Post-seizure observation correlations.}
% \label{fig:SpatialCrossCorrelation}
% \end{figure}
% 
% \begin{figure}[!ht]
% \begin{center}
% \includegraphics{./Figures/HomoTestCrossCorr.eps}
% \end{center}
% \caption{{\bf Mean cross-correlations between consecutive samples using subsets of observations of the neural field}. Each subplot shows correlations corresponding to a subset electrodes. Indexing the electrodes in a (x,y) grid, where index ranges from 1 to 10. Subplot A shows the mean correlation between electrode indexes (1-6, 5-10), subplot B shows correlations from indexes (5-10,5-10), subplot C shows indexes (1-6, 1-6) and subplot D shows indexes (5-10, 5-10).}
% \label{fig:HomogeneityTest}
% \end{figure}
